Using simulation to optimize the sampling design of fisheries-independent trawl surveys

### 2019-03-07 analysis/paper/sim_survey_paper.Rmd
Complete draft of new results section

The SimSurvey package serves as a tool for simulating surveys of dynamic populations that vary across ages, time and space. The core of the simulation is based on the widely used cohort equation and, even though the processes that define recruitment and total mortality are simple, a wide range of stock dynamics can be simulated by changing a few parameters (Figure \@ref(fig:surface)). This base population can then be distributed through a grid (Figure \@ref(fig:grid)) and relationships with depth can be defined as can the nature of the correlation across ages, years and space (Figure \@ref(fig:distribution)). Together, two functions (`sim_abundance` and `sim_distribution`) are capable of simulating a wide range of populations with different life histories and spatial properties. The next necessary step to generating data similar to actual observations is to conduct a survey. In this package we implement a function, `sim_survey`, that conducts a stratified random survey of the population. The sampling process is governed by the area covered by the trawl as well as age-specific catchability. Sub-sampling protocol (length and age sampling) can also be varied. As such, data from a wide range of surveys can be simulated (Figure \@ref(fig:survey)).

- Make the results more general, almost about the flexibility of the package, then hone in on the case study??

A dynamic population that varies across ages, time and space was simulated with the parameter settings used here (Table \@ref(tab:par)). Because the default settings of the functions match the case study settings, Figure \@ref(fig:surface)a depicts abundance at age through time and Figure \@ref(fig:distribution)a depicts the spatial distribution of ages 1-3 through years 1-3. In general, the parameter settings used dictate that all ages tend to aggregate in fairly dense clusters and that low-density zones are relatively wide-ranging. Surveys, therefore, tend to be characterized by regions with large catches interspersed with regions of no catch (Figure \@ref(fig:survey)). The samples within each successful set are also dynamic and rarely include all age groups. This is because only moderate correlation ($\varphi_{\xi, age}$ of 0.5) was imposed on the distributions of ages 1-5+, allowing individuals of different ages to occupy different locations. These settings were chosen to simulated data that roughly correspond to actual survey data of cod from NAFO division 3Ps, where cod between the ages of 1-5 tend to be caught at different locations.

The stratified analyses run on these simulated data provided estimates of total abundance, abundance at length and abundance at age, and different stages of sampling affect these estimates. Specifically, set density affects all estimates, length sampling affects estimates of abundance at length and abundance at age, and age sampling affects abundance at age estimates. This is because stratified estimates of total abundance is based on aggregate catch while abundance at length estimates requires the length frequency data and abundance at age estimates requires both length and age data to construct an age-length key.

Figure \@ref(fig:tot) depicts the consequences of different set densities on the estimates of total abundance available to the survey through the simulated series. It is clear from this figure that the probability envelopes are tighter (i.e. estimates are more precise) and bias is negligible. Likewise, abundance at length and abundance at age estimates are more precise as set density is increased (Figure \@ref(fig:length), \@ref(fig:age), \@ref(fig:year)). With the exception of a length sampling protocol of 10 measurements per set ($M_{lengths}$ = 10), there appears to be little difference in the estimates of abundance at length when the number of measurements are capped at 100 ($M_{lengths}$ = 100) or 1000 ($M_{lengths}$ = 1000; Figure \@ref(fig:length)). Some downward bias is apparent in the abundance at length estimates, especially at larger lengths, under the low to medium set density scenarios with low length sampling effort ($D_{sets}$ = 0.0005 or 0.002 and $M_{lengths}$ = 10; Figure \@ref(fig:length)). Upward or downward bias in abundance at age estimates is apparent in all cases except the high set density and high length sampling scenario ($D_{sets}$ = 0.01 and $M_{lengths}$ = 1000; Figure \@ref(fig:age), \@ref(fig:year)). Under both low and medium set densities ($D_{sets}$ = 0.0005 and 0.02, respectively), increasing length sampling effort from 10 or 100 measurements per set to 1000 appears to increase bias or reduce precision slightly (Figure \@ref(fig:age), \@ref(fig:year)).

For a less visual approach to assessing precision and bias, RMSE results are presented in Figures \@ref(fig:error-lines) and \@ref(fig:error-wireframe) and Table \@ref(tab:comp). Figure \@ref(fig:error-lines) depicts the effects of set density and length sampling effort on abundance at length and abundance at age estimates. In all cases, increases in set density results in large declines in RMSE (Figure \@ref(fig:error-lines)). For abundance at length estimates, RMSE declines as the number of lengths measured increases, however, gains plateau beyond a cretin point. Likewise, RMSE of abundance at age estimates decline with increased length sampling effort, but RMSE does not plateau in all set density scenarios (Figure \@ref(fig:error-lines)). While RMSE continues to decline with increased length sampling effort under the high set density scenarios ($D_{sets}$ = 0.005 to 0.01 sets / km^2^), RMSE reaches a plateau and then increases with increased length sampling effort under the medium to low set density scenarios ($D_{sets}$ = 0.002 to 0.0005 sets / km^2^; Figure \@ref(fig:error-lines)). In many cases, RMSE for both metrics is lower for scenarios with higher set density but lower length sampling effort (Figure \@ref(fig:error-lines)). Similar patterns are apparent in Figure \@ref(fig:error-wireframe) and Table \@ref(tab:comp), which focus on the consequences of set density and length and age sampling protocol on the estimates of abundance at age. Again, RMSE clearly declines with increases to set density and sometimes declines with increased length sampling effort (Figure \@ref(fig:error-wireframe)). Under both low and medium set density scenarios ($D_{sets}$ = 0.005 to 0.01 sets / km^2^), measuring a maximum of 100 fish per set appears to be better than measuring a maximum of 1000 fish per set (Figure \@ref(fig:error-wireframe); Table \@ref(tab:comp)). Decreasing the age sampling protocol below 10 ages sampled per set per length group appears to result in relatively small increases to RMSE (Figure \@ref(fig:error-wireframe); Table \@ref(tab:comp)).

>>>>

The SimSurvey package serves as a tool for simulating stratified random surveys of dynamic populations that vary across ages, time and space. The core of the simulation is based on the widely used cohort equation and, even though the processes that define recruitment and total mortality are simple, a wide range of stock dynamics can be simulated by changing a few parameters (Figure \@ref(fig:surface)). This base population can then be distributed through a grid (Figure \@ref(fig:grid)) and relationships with depth can be defined as can the nature of the correlation across ages, years and space (Figure \@ref(fig:distribution)). Together, two functions (`sim_abundance` and `sim_distribution`) are capable of simulating a wide range of populations with different life histories, depth associations and spatial properties. The next necessary step to generating data similar to actual observations is to conduct a survey. In this package we implement a function, `sim_survey`, that conducts a stratified random survey of the population. The sampling process is governed by the area covered by the trawl as well as age-specific catchability. Sub-sampling protocol (length and age sampling) can also be varied. As such, data from a wide range of surveys can be simulated (Figure \@ref(fig:survey)). There are a plethora of statistical models that may be tested using these simulated data, however, implementing a variety of analylitical approaches was outside the scope of this work. Insted we focus on analyzing simulated stratified-random survey data using a design-based stratified analysis. A stratified analyis is facilitated using the `run_strat` function and the precision and accuracy of the results (e.g. RMSE) can be calculated using `strat_error`. This is a simple and widly-used analysis, and the speed at which it runs allows for a wide range of survey designs to be tested, via the `test_surveys` function, in a reasonable timeframe.

Results from the case study presented provide one example of how this package can be used to explore the effacacy of various survey designs. A dynamic population with a patchy and age-clustered distribution was simulated for this case study (Figure \@ref(fig:surface)a, \ref(fig:distribution)a). In general, the parameter settings used (function defaults) dictate that all ages tend to aggregate in fairly dense clusters and that low-density zones are relatively wide-ranging. Surveys, therefore, tend to be characterized by regions with large catches interspersed with regions of no catch (Figure \@ref(fig:survey)). The samples within each successful set are also dynamic and rarely include all age groups. This is because only moderate correlation ($\varphi_{\xi, age}$ of 0.5) was imposed on the distributions of ages 1-5+, allowing individuals of different ages to occupy different locations. These settings were chosen to simulated data that roughly correspond to actual survey data of cod from NAFO division 3Ps, where cod between the ages of 1-5 tend to be caught at different locations.

A series of surveys were run over the simulated population (see `expand_surveys` call in Table \@ref(tab:par)) and stratified analyses were run on the data obtained. The stratified analyses provided estimates of total abundance, abundance at length and abundance at age, and different stages of sampling affect these estimates. Specifically, set density affects all estimates, length sampling affects estimates of abundance at length and abundance at age, and age sampling affects abundance at age estimates. This is because stratified estimates of total abundance is based on aggregate catch while abundance at length estimates requires both total and length frequency data; finally, abundance at age estimates requires set totals, and length and age data to construct an age-length key.

The effect of set density on estimates of total abundance, abundance at length and abundance at age are clear across all figures generated from the `test_surveys` results (Figures \@ref(fig:fan1), \@ref(fig:fan2), \@ref(fig:fan3), \@ref(fig:rank), \@ref(fig:rmse-surface)). Across all fan plots, it is clear that the probability envelopes tighten (i.e. estimates are more precise) as set density is increased (Figures \@ref(fig:fan1), \@ref(fig:fan2), \@ref(fig:fan3)). Clear declines in RMSE are also apparant in estimates of abundance at length as set density is increased (Figure \@ref(fig:rank)) and there are notable differences in the scale of RMSE in estimates of abundance at age as set density is increased (Figure \@ref(fig:rmse-surface)).

Compared to increases in set density, the effects of increased length sampling effort on abundance at length estimates are less clear. Regarding estimates of abundance at length, RMSE appears to reach a platau at when length sampling effort is increased 100 measurements per set (Figure \@ref(fig:rank)). Increasing the sampling rule above 100 measurements per set results in many more fish being measured, however, the increase in sampling effort is not matched with substantive declines in RMSE (Figure \@ref(fig:rank)). Interestingly, it appears that measuring fewer total fish at more locations appears to be more benificial than many fish at fewer locations (Figure \@ref(fig:rank)).

All levels of sampling affect the abundance at age estimates as set catches define the maginitude of the population, length sampling define the length distribution and data from the length-stratified age sampling is used to construct an age-length-key to convert length frequenceis to age frequencies. Like the abundance at length estimates, the greatest improvements to RMSE come from increasing set density rather than sub-sampling effort (Figure \@ref(fig:rmse-surface)). Specifically, decreasing the age sampling protocol below 10 ages sampled per length group per division appears to result in relatively small increases to RMSE (Figure \@ref(fig:rmse-surface)). Lenght sampling effort, in contrast, has an uneven impact depending on the set density scenario. At low set densities ($D_{sets}$ = 0.0005 sets / km^2^), RMSE declines when length sampling effort is increased from around 5 to around 100 measurements per set, and RMSE starts to increase as length sampling effort is increased beyond ~100 measurements per set; in fact, RMSE values appear higher at the highest length sampling scenario (1000 measurements / set) than the lowest (5 measurements / set; Figure \@ref(fig:rmse-surface)). A similar pattern is apparant under the medium set density scenario ($D_{sets}$ = 0.002 sets / km^2^), however, RMSE values under the lowest and highest length sampling scenarios are of similar magnitude. Finally, RMSE continues to decline with increased length sampling effort under the high set density scenarios ($D_{sets}$ = 0.01 sets / km^2^; Figure \@ref(fig:rmse-surface)).
